Phylogenetics and the human microbiome.

### 2014-07-18 sbreview.tex
Smoothing out the local sentences.

To calculate the UniFrac (``unique fraction'') divergence between two samples of reads, one builds a global tree based on both samples and then calculates the fraction of the total tree length found in only one (i.e. the fraction of the tree unique to one) of the two samples.
Specifically, the original UniFrac divergence is the fraction of the total branch length of the tree coming from branches with descendant leaves in exactly one of the two samples \citep[Fig.~\ref{fig:unifrac};][]{LozuponeKnightUniFrac05}.
Weighted UniFrac is an abundance weighted version \citep{LozuponeEaWeightedUnifrac07}.
These dissimilarity measures have hundreds of citations.
\citet{evans2012phylogenetic} showed that weighted UniFrac is in fact a specific case of the earth-movers distance, and that the commonly-used randomization procedure for significance estimation has a central limit theorem approximation.
The earth-movers distance \citep{monge1781memoire,Villani2003-wv} between two probability distributions in this case can be defined using a physical analogy as the minimum amount of ``work'', defined as mass times distance, required to move probability mass in one distribution to another along the tree.
In this case the size of the dirt piles is proportional to the number of reads mapping to that location in the tree (Fig.~\ref{fig:dirtpiles}).
\citet{chen2012associating} have shown that a partially abundance-weighted variant of UniFrac may have greater power to resolve community differences than either unweighted or weighted UniFrac.
There are clear connections between UniFrac and phylogenetic diversity explored by \citet{faith2009cladistic}, who point out related measures in Faith's earlier work.

>>>>

To calculate the unweighted UniFrac (``unique fraction'') divergence between two samples of reads, one builds a global tree based on both samples and then calculates the fraction of the total tree length found in only one (i.e. the fraction of the tree unique to one) of the two samples \citep[Fig.~\ref{fig:unifrac};][]{LozuponeKnightUniFrac05}.
Weighted UniFrac is an abundance weighted version \citep{LozuponeEaWeightedUnifrac07}.
These dissimilarity measures have hundreds of citations.
\citet{evans2012phylogenetic} showed that weighted UniFrac is in fact a specific case of the earth-movers distance, and that the commonly-used randomization procedure for significance estimation has a central limit theorem approximation.
The earth-movers distance \citep{monge1781memoire,Villani2003-wv} between two probability distributions in this case can be defined using a physical analogy as the minimum amount of ``work'', defined as mass times distance, required to move probability mass in one distribution to another along the tree.
In this case the size of the dirt piles is proportional to the number of reads mapping to that location in the tree (Fig.~\ref{fig:dirtpiles}).
\citet{chen2012associating} have shown that a partially abundance-weighted variant of UniFrac may have greater power to resolve community differences than either unweighted or weighted UniFrac.
There are clear connections between UniFrac and phylogenetic diversity explored by \citet{faith2009cladistic}, who point out related measures in Faith's earlier work.
